Global Optimization of the Analogue Method by Means of Genetic Algorithms

### 2015-11-20 Genetic algorithms - Journal of Climate/template.tex
Michel's corrections.
Removing we/our in the article on GAs.

\abstract{The analogues method is based on a statistical relationship between synoptic atmospheric variables (predictors) and local weather (predictand), which we aim at forecasting. This relationship is expressed through many parameters that are usually calibrated by means of a semi-automatic procedure. This sequential calibration has strong limitations: it is made of successive steps and thus cannot handle parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely genetic algorithms, which can optimize jointly all parameters of the method and get closer to a global optimum by taking into account the parameters dependencies. Moreover, it can choose objectively parameters that were previously manually assessed, and can take into account new degrees of freedom that were unthinkable before.
These kind of optimization techniques need however to be tailored to the problem to solve. We had to assess multiple combinations of algorithms, and even to develop new operators, such as the \textit{chromosome of adaptive search radius} which was found to be very robust, in order to make recommendations for the use of genetic algorithms to optimize the analogue method. These recommendations are a consequent outcome of this work, as it opens new perspective for the improvement of the analogue method, and its application to new regions or to new predictands.
}

@@

We observed during the AM calibration that the resulting parameters vary with the initial choices (such as the number of analogues). In addition, the different levels of analogy (eg. on the atmospheric circulation and the moisture index) are always calibrated sequentially. However, we cannot exclude any dependency between them, which could lead us to select another parametrization if we calibrate them together. Simultaneous calibration of all parameters has never been undertaken so far. Only a global optimization technique can be able to optimize all parameters of all analogy levels simultaneously.

@@

Our ambition was then to assess the feasibility of an automatic optimization of the AM in order to assess all parameters simultaneously, and thus to be able to get nearer to the global optimum in the parameters space. In addition, it can overcome the systematic manual assessments of all pressure levels and temporal windows. Finally, it can open new perspectives by allowing the addition of new degrees of freedom, such as a weighting of the criteria values between the pressure levels \cite[see]{Horton2016b}, and the consideration of non-overlapping spatial windows between the pressure levels.

@@

The GAs optimize a population of $N$ individuals (parameter sets). Each individual contains a chromosome (parameters of the AM in our case). We call gene every parameter that constitutes the chromosome. The floating-point representation \citep[instead of the canonical binary, see][]{Goldberg1989, Goldberg1990a, Holland1992b} of the genes was considered, as it was found more suited in multiple applications \citep{Michalewicz1996, Herrera1998a, Haupt2004, Back1996b, Gaffney2010a}.

@@

All operators we used and their options, applied to floating-point coding, are described in the following sections. Many other operators exist, but we will only present the ones we evaluated.

@@

The first step of the optimization is to generate the initial population. A population is a set of $N$ individuals (each of which represents a point in the space of potential solutions, a parameter set of the AM in our application) that we are going to make evolve. A generation is the population at a given time.

@@

	\textit{Roulette wheel weighting on fitness}: the selection probability is calculated based on the value of the objective function. The probability $p_{n}$ of each individual is calculated by the equation \ref{equation_mating_score_weighting}:
	\begin{equation}
	p_{n}=\frac{score_{n}-score_{N_{mp}}}{\sum_{n=1}^{N_{mp}} (score_{n}-score_{N_{mp}})}
	\label{equation_mating_score_weighting}
	\end{equation}
	In our application, the last individual ($N_{mp}$) has zero probability of being selected.

@@

\begin{itemize}
	\item Single-point crossover: a crossover point is randomly chosen for the pair. The genes (our parameters) located after that point are exchanged in between the two chromosomes.

@@

	\item Variable uniform mutation \citep{Fogarty1989}: a variable mutation rate over the generations was first suggested by \citet{Holland1992b} and evaluated by \citet{Fogarty1989}. In most applications, the mutation rate decreases with the generations, in a deterministic and global (for all individuals) manner \citep{Back1992b}. Its optimum configuration depends on the size of the chromosomes, of the properties of the objective function, and of the population size \citep{Back1992b}. We implemented this operator according to equation \ref{equation_mutation_uniformvariable}.
	\begin{equation}
	p_{n,G} = p_{G_{0}}+\left( \dfrac{p_{G_{0}}-p_{G_{m,p}}}{G_{m,p}} \right) min\left\lbrace G,G_{m,p}\right\rbrace 
	\label{equation_mutation_uniformvariable}
	\end{equation}
	where $p_{n,G}$ is the mutation rate (probability) of the gene $n$ for generation number $G$, $G_{m,p}$ is the maximum number of generations during which the mutation rate varies. $p_{G_{0}}$ is the initial mutation probability, and $p_{G_{m,p}}$ is the final one. $p_{G_{0}}$, $p_{G_{m,p}}$ and $G_{m,p}$ are the three controlling parameters of the operator. The evolution of the mutation rate is linear.

@@

	\item Variable normal mutation: with the same logic as the variable uniform mutation, we tested a mutation operator using a normal distribution with a variable mutation rate and standard deviation. The mutation rate is calculated with equation \ref{equation_mutation_uniformvariable}. On the same principle, we decrease linearly the standard deviation over the generations:
	\begin{equation}
	\sigma_{n,G} = \sigma_{G_{0}}+\left( \dfrac{\sigma_{G_{0}}-\sigma_{G_{m,\sigma}}}{G_{m,\sigma}} \right) min\left\lbrace G,G_{m,\sigma}\right\rbrace 
	\label{equation_mutation_normalvariable}
	\end{equation}
	where $\sigma_{n,G}$ is the standard deviation of gene $n$ and generation number $G$, $\sigma_{G_{0}}$ is the initial standard deviation, $\sigma_{G_{m,\sigma}}$ is the final standard deviation, $G_{m,\sigma}$ is the maximum number of generations during which the standard deviation varies. $p_{G_{0}}$, $p_{G_{m,p}}$, $G_{m,p}$, $\sigma_{G_{0}}$, $\sigma_{G_{m,\sigma}}$ and $G_{m,\sigma}$ are the six parameters of the method.

@@

	We adapted this operator for our application, which is not based on a predefined number of generations:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} \varphi^{2} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} \varphi^{2} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_nonuniform}
	\end{equation}

@@

	\item Individual adaptive search radius (new): based on the ideas of the non-uniform mutation, we introduce a search radius in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between 0 and 1, is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is considered initially as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_rayon_adaptatif}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, in the same way as for the non-uniform mutation. No external parameter is therefore necessary.

	\item Chromosome of adaptive mutation rate \citep[or \textit{n adaptative mutation rate},][]{Back1992a}: analogously to the individual adaptive mutation rate, this approach leaves the control of the evolution rate to the individuals themselves. The difference here is that each gene has a specific mutation rate. The main advantage is that the tuning of the mutation can be much more precise \citep{Smith1997a}. We therefore consider a second chromosome containing the mutation rate for each gene of the first chromosome. The operations of mutation and self-mutation are similar to the case of the individual adaptive mutation rate, but in a distributed way, within the chromosome. Another difference is that we apply the same crossover operations as those applied to the first chromosome, and this for the same crossing points. Thus, during an exchange of genes, children also inherit the mutation rates specific for each of these genes.

	\item Chromosome of adaptive search radius (new): we introduced this operator that combines the operations of the chromosome of adaptive mutation rate to our adaptive search radius approach. Similarly, an individual has 3 chromosomes: the first containing the values to be optimized, the second the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameters are required.

	\item Multi-scale mutation (new): finally, we developed another approach, that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. We wanted to test an approach that combines both exploration and exploitation during the whole optimization. Thus, we considered the search radius $r_{a}$ of equation \ref{equation_mutation_rayon_adaptatif} as a random value for each individual, but restricted to 4 equiprobable values: 1, 0.5, 0.1, 0.02, which range from full exploration to fine exploitation. The only external parameter is the mutation rate which is fixed.

@@

We used a process of elitism on the natural selection as well as on mutations. This ensures the survival of the best individual so that we do not lose a better solution. After the natural selection operator, if the best individual has not been selected, it is copied to the IG instead of an individual randomly picked. After mutation, if the best individual has mutated and if its new version has a lower performance score than the original, the latter is also reinserted in the IG instead of an individual randomly chosen.

@@

The convergence check determines whether the solution is acceptable and if the algorithm may stop. The stopping criteria are not often well documented in GAs case studies. We chose to stop the optimization if the best individual does not change for $x$ generations. This value should not be too low to allow the algorithm to escape from a local optima. In addition, the rate of improvement decreases with the progression of the optimization. It is thus common that the best individual does not evolve over several generations when we get closer to the global solution. A value of $x=20$ generations was chosen.

@@

GAs are very computationally intensive because they require many evaluations of the objective function. These assessments are very long in our application, as they require calculating and assessing a forecast for every day of the calibration period, thus over several decades. In order to reduce the computation time, we avoid recalculating the performance score of an individual who has previously been evaluated and that has not changed. We keep the score of each individual living in the selection until it mutates.

@@

One of our goals being to make recommendations of parametrization in order to optimize the AM, we proceeded systematically. The results are summarized hereafter \citep[see][for the details]{Horton2012a}. We used concepts from the factorial design approach \citep[see eg.][]{Costa2005a,Costa2007a,Mariano2010a}, which is sometimes used for comparative analysis of different parametrizations of GAs. It allows isolating the effect of a parameter under different combinations of the other options. We processed by stages, analysing in details and in a systematic way every variants of the implemented operators, in combination with multiple other options and parameters in order to take into account eventual co-dependencies. Our goal here is not to focus on the performance score we obtain through optimization, neither the values of the new optimized parameters \citep[covered in][]{Horton2016b}, but to explain how to use GAs to optimize AMs in an efficient way.

In order to evaluate a combination of operators/options, we processed 10 optimizations per parametrization of GAs. The performances were characterized by four indicators: (i) mean performance score: average of the final scores of the 10 optimizations, (ii) convergence: the number of optimizations that converged to a supposed global optimum, (iii) number of generations: characterization of the convergence speed, and (iv) number of evaluations of the objective function: characterization of the required calculation time.

@@

After a first overview of the results, GAs have quicky been proven successful at optimizing the AM. Indeed, the performance score of the sequential approach has been quickly exceeded without adding any new parameter to the method (Figure \ref{fig:gas_evolution_good}). Even GAs parametrization that we will consider further as inadequate (Figure \ref{fig:gas_evolution_bad}) did significantly better that the sequential approach.

@@

The results illustrate the effect of an operator when its contribution is isolated from the other operators. It means that we compare the effect of the studied operator when all the other operators are equivalent, while assessing multiple combinations. We then summarize this contribution relatively to the other options, as a percentage of gain/loss regarding the mean of all variants. For example, to evaluate the performance of the uniform crossover operator, we compare its performance to the average of all crossover operators while retaining the same population size, the same mutation operators, natural selection, and selection of couples.

@@

We evaluated every combination of 6 options for the couples selection (Table \ref{tab:assessed_couples_selection_operators}) and 21 for the chromosome crossover operators (Table \ref{tab:assessed_crossover_operators}), along with variants of the other operators. This results in 1,008 combinations, requiring 10,080 optimizations.

@@

The performance of the couples selection operator are relatively close (Figure \ref{fig:operator_selectcoupl_score}). Overall, we observe that the tournament selection with 3~candidates is slightly superior to others, along with the roulette wheel weighting. This last one is however a bit less effective in terms of convergence and number of evaluations (not shown). The couples selection operator has not a significant role in our application.

@@

Having identified the leading role of the mutation operator, the sensitivity analysis focused on it. Each of the 10 different implementations (see section \ref{sec:gas:mutation}) was tested, with different parameters for those who require some (Table \ref{tab:assessed_mutation_operators}), bringing the number of its variations up to 109. We also did some optimizations without any mutation as a reference. Along with variants of the other operators \citep[see][for the details]{Horton2012a}, it resulted in 660 combinations (so 6,600 optimizations).

@@

The mutation operators based on a variable normal or variable uniform laws work very poorly and are difficult to configure. We then observe many operators more or less with the same performance scores and requiring a variable amount of assessments. The convergence analysis \citep[see][]{Horton2012a} allows us to highlight three best operators: non-uniform mutation, chromosome of adaptive search radius, and multi-scale mutation. Because of the importance of the mutation operator on the resulting performance, we studied its role and parametrization in more detail. Thus, we performed different optimizations using variants of these 3 operators (Table \ref{tab:assessed_mutation_operators_bis}).

@@

The first analysis was the optimization of the precipitation forecasting over a subcatchment (Binn-Simplon region) in the Swiss Alps (Figure \ref{fig:operator_mutation_score_atmlevel}). The optimizer could choose the 2 pressure levels of the atmospheric circulation analogy (method with a single level of analogy). The resulting performance score \citep[CRPSS = skill score of the CRPS, see][]{Brown1974} is obviously superior to the one from the sequential calibration (red line on Figure \ref{fig:operator_mutation_score_atmlevel}). For most options, it is also slightly better than the results from the optimization without selection of the pressure levels (blue line). A clear breakthrough of the performances was not expected, as the former selection of pressure levels results from intensive work \citep{Bontron2004}. This application however demonstrates that, when correctly parametrized, GAs can automatically and successfully choose the pressure levels. However, some less relevant optimizations do not match the previous results. We could clearly observe, through different applications, that the automatic selection of the pressure level significantly increases the difficulty for GAs to converge to a unique solution, ideally the global optimum. A difficulty is that the pressure levels are considered within the optimization without continuity between values, and thus the approaches relying on distance in the parameters space, such as the search radius, cannot fully exploit the properties that make them efficient. However, even though the results show a certain variability, most of them present very good performance scores, despite different parameters of the AM.

Then, we performed the same optimization, but for another region, sensitive to other meteorological influences (Figure \ref{fig:operator_mutation_score_rhoneamont}), in order to assess eventual dependencies of the operator with the predictand. Even though we can observe differences with Figure \ref{fig:operator_mutation_score_atmlevel}, it is globally the same options that perform better.

Next, we added a second level of analogy (Figure \ref{fig:operator_mutation_score_r2}) on moisture variables \citep[see][]{Bontron2004, Horton2016}. GAs had to optimize both levels of analogy simultaneously. Once again, despite the difficulty to optimize simultaneously both levels of analogy, the results were better than the sequential calibration (red line on Figure \ref{fig:operator_mutation_score_r2}). And finally, we added a preselection on air temperature instead of the fixed calendar window, as proposed by \cite{BenDaoud2015}. The results show generally higher scores (Figure \ref{fig:operator_mutation_score_r4}), demonstrating the success of the optimizer to take advantage of this new degree of freedom, and its capacity to handle optimization of 3 analogy levels jointly. Again, the most relevant options are generally the same.

After analysis of the most relevant mutation operators, we can raise the following advice (detailed parametrization are provided in section \ref{sec:recommendations}):

@@

	\item \textit{Non-uniform mutation} \citep{Michalewicz1996}: this operator is good in terms of convergence, mainly when the number of parameters to optimize is rather low. The number of required evaluations, however, can be quite substantial. The main disadvantage of the non-uniform mutation is the number of parameters it requires, that are difficult to estimate a priori. We could observe that the $\omega$ coefficient does not influence performance. The role of $G_{m}$ is rather difficult to judge, but does not seem essential. The mutation rate was found to be important. The difficulty is that the optimal value may be case-related.

	\item \textit{Chromosome of adaptive search radius} (new): unlike the previous one, our new operator is very robust, as it requires no option and is auto-adapting. It is interesting to notice that our insertion of an extra chromosome representing the search radius gives better performance than other self-adaptive operators (such as, for example, the chromosome of adaptive mutation rate). If one had to choose a single option for the mutation operator, we would recommend this one, as it was proven effective and needs no parameter.

@@

In our application, the mutation operator has a leading effect and should be chosen with care. It may be wise to perform multiple optimizations and to consider these three operators in parallel in order to obtain results from options that are either sometimes more efficient or more robust. It is interesting to note that the three best techniques incorporate a notion of search distance. It is likely that this notion is the key to these algorithms, for our application, and allows them to initially explore the parameter domain, and then to converge. The search radius in fact directly represents the notion of transition between exploration and exploitation, in our opinion more than a possible evolution of mutation rates.

@@

We also assessed the influence of the size of the IG (ratio of the total population) selected for mating (Figure \ref{fig:option_popratio_score}). It does not appear that this parameter is critical to the quality of the optimizations, provided it is not too big. A value of 50~\% seems a wise choice.

@@

We evaluated the optimization by GAs on AMs of varying complexity, with a large number of combinations of operators to be able to make recommendations for optimizing the AM. Our conclusions are:

@@

The optimization does not systematically converge to the global optimum  (but still often nearby), which is why we recommend doing several optimizations in parallel in order to compare the results, analyze the convergence, and keep the best. It may be wise to consider the three mutation operators in parallel. It would then combine algorithms that are sometimes faster and other that are more robust. In order to be confident in the optimized AMs, we propose using a set of the following mutation operators:

>>>>

\abstract{The analogues method is based on a statistical relationship between synoptic atmospheric variables (predictors) and local weather (predictand), which we aim at forecasting. This relationship is expressed through many parameters that are usually calibrated by means of a semi-automatic procedure. This sequential calibration has strong limitations: it is made of successive steps and thus cannot handle parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely genetic algorithms, which can optimize jointly all parameters of the method and get closer to a global optimum by taking into account the parameters dependencies. Moreover, it can choose objectively parameters that were previously manually assessed, and can take into account new degrees of freedom that were unthinkable before.
These kind of optimization techniques need however to be tailored to the problem to solve. Multiple combinations of algorithms had to be assessed, and even new operators were developed, such as the \textit{chromosome of adaptive search radius} which was found to be very robust, in order to make recommendations for the use of genetic algorithms to optimize the analogue method. These recommendations are a consequent outcome of this work, as it opens new perspective for the improvement of the analogue method, and its application to new regions or to new predictands.
}

@@

During the AM calibration it has been found that the resulting parameters may vary with initial choices (such as the number of analogues). In addition, the different levels of analogy (eg. on the atmospheric circulation and the moisture index) are always calibrated sequentially. However, any dependency between them cannot be excluded, which could lead us to select another parametrization if calibrated together. Simultaneous calibration of all parameters has never been undertaken so far. Only a global optimization technique can be able to optimize all parameters of all analogy levels simultaneously.

@@

The goal was then to assess the feasibility of an automatic optimization of the AM in order to assess all parameters simultaneously, and thus to be able to get nearer to the global optimum in the parameters space. In addition, it can overcome the systematic manual assessments of all pressure levels and temporal windows. Finally, it can open new perspectives by allowing the addition of new degrees of freedom, such as a weighting of the criteria values between the pressure levels \cite[see]{Horton2016b}, and the consideration of non-overlapping spatial windows between the pressure levels.

@@

The GAs optimize a population of $N$ individuals (parameter sets). Each individual contains a chromosome (parameters of the AM in this case). Genes are all parameters constituting the chromosome. The floating-point representation \citep[instead of the canonical binary, see][]{Goldberg1989, Goldberg1990a, Holland1992b} of the genes was considered, as it was found more suited in multiple applications \citep{Michalewicz1996, Herrera1998a, Haupt2004, Back1996b, Gaffney2010a}.

@@

All considered operators and their options, applied to floating-point coding, are described in the following sections. Many other operators exist, but only the ones evaluated are presented.

@@

The first step of the optimization is to generate the initial population. A population is a set of $N$ individuals (each of which represents a point in the space of potential solutions, a parameter set of the AM in this application) that are going to evolve. A generation is the population at a given time.

@@

	\textit{Roulette wheel weighting on fitness}: the selection probability is calculated based on the value of the objective function. The probability $p_{n}$ of each individual is calculated by the equation \ref{equation_mating_score_weighting}:
	\begin{equation}
	p_{n}=\frac{score_{n}-score_{N_{mp}}}{\sum_{n=1}^{N_{mp}} (score_{n}-score_{N_{mp}})}
	\label{equation_mating_score_weighting}
	\end{equation}
	In this application, the last individual ($N_{mp}$) has zero probability of being selected.

@@

\begin{itemize}
	\item Single-point crossover: a crossover point is randomly chosen for the pair. The genes (AM parameters) located after that point are exchanged in between the two chromosomes.

@@

	\item Variable uniform mutation \citep{Fogarty1989}: a variable mutation rate over the generations was first suggested by \citet{Holland1992b} and evaluated by \citet{Fogarty1989}. In most applications, the mutation rate decreases with the generations, in a deterministic and global (for all individuals) manner \citep{Back1992b}. Its optimum configuration depends on the size of the chromosomes, of the properties of the objective function, and of the population size \citep{Back1992b}. This operator was implemented according to equation \ref{equation_mutation_uniformvariable}.
	\begin{equation}
	p_{n,G} = p_{G_{0}}+\left( \dfrac{p_{G_{0}}-p_{G_{m,p}}}{G_{m,p}} \right) min\left\lbrace G,G_{m,p}\right\rbrace 
	\label{equation_mutation_uniformvariable}
	\end{equation}
	where $p_{n,G}$ is the mutation rate (probability) of the gene $n$ for generation number $G$, $G_{m,p}$ is the maximum number of generations during which the mutation rate varies. $p_{G_{0}}$ is the initial mutation probability, and $p_{G_{m,p}}$ is the final one. $p_{G_{0}}$, $p_{G_{m,p}}$ and $G_{m,p}$ are the three controlling parameters of the operator. The evolution of the mutation rate is linear.

@@

	\item Variable normal mutation: with the same logic as the variable uniform mutation, a mutation operator was tested using a normal distribution with a variable mutation rate and standard deviation. The mutation rate is calculated with equation \ref{equation_mutation_uniformvariable}. On the same principle, the standard deviation decreases linearly over the generations:
	\begin{equation}
	\sigma_{n,G} = \sigma_{G_{0}}+\left( \dfrac{\sigma_{G_{0}}-\sigma_{G_{m,\sigma}}}{G_{m,\sigma}} \right) min\left\lbrace G,G_{m,\sigma}\right\rbrace 
	\label{equation_mutation_normalvariable}
	\end{equation}
	where $\sigma_{n,G}$ is the standard deviation of gene $n$ and generation number $G$, $\sigma_{G_{0}}$ is the initial standard deviation, $\sigma_{G_{m,\sigma}}$ is the final standard deviation, $G_{m,\sigma}$ is the maximum number of generations during which the standard deviation varies. $p_{G_{0}}$, $p_{G_{m,p}}$, $G_{m,p}$, $\sigma_{G_{0}}$, $\sigma_{G_{m,\sigma}}$ and $G_{m,\sigma}$ are the six parameters of the method.

@@

	This operator was adapted for this application, which is not based on a predefined number of generations:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} \varphi^{2} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} \varphi^{2} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_nonuniform}
	\end{equation}

@@

	\item Individual adaptive search radius (new): based on the ideas of the non-uniform mutation, a search radius was introduced in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between 0 and 1, is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is considered initially as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_rayon_adaptatif}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, in the same way as for the non-uniform mutation. No external parameter is therefore necessary.

	\item Chromosome of adaptive mutation rate \citep[or \textit{n adaptative mutation rate},][]{Back1992a}: analogously to the individual adaptive mutation rate, this approach leaves the control of the evolution rate to the individuals themselves. The difference here is that each gene has a specific mutation rate. The main advantage is that the tuning of the mutation can be much more precise \citep{Smith1997a}. A second chromosome containing the mutation rate for each gene of the first chromosome was therefore considered. The operations of mutation and self-mutation are similar to the case of the individual adaptive mutation rate, but in a distributed way, within the chromosome. Another difference is that the same crossover operations are applied as those applied to the first chromosome, and this for the same crossing points. Thus, during an exchange of genes, children also inherit the mutation rates specific for each of these genes.

	\item Chromosome of adaptive search radius (new): this operator combines the operations of the chromosome of adaptive mutation rate to the adaptive search radius approach. Similarly, an individual has 3 chromosomes: the first containing the values to be optimized, the second the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameters are required.

	\item Multi-scale mutation (new): finally, another approach was developed that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. The idea was to test an approach that combines both exploration and exploitation during the whole optimization. Thus, the search radius $r_{a}$ of equation \ref{equation_mutation_rayon_adaptatif} was considered as a random value for each individual, but restricted to 4 equiprobable values: 1, 0.5, 0.1, 0.02, which range from full exploration to fine exploitation. The only external parameter is the mutation rate which is fixed.

@@

A process of elitism was introduced on the natural selection as well as on mutations. This ensures the survival of the best individual so that a better solution is never lost. After the natural selection operator, if the best individual has not been selected, it is copied to the IG instead of an individual randomly picked. After mutation, if the best individual has mutated and if its new version has a lower performance score than the original, the latter is also reinserted in the IG instead of an individual randomly chosen.

@@

The convergence check determines whether the solution is acceptable and if the algorithm may stop. The stopping criteria are not often well documented in GAs case studies. The optimization is here stopped if the best individual does not change for $x$ generations. This value should not be too low to allow the algorithm to escape from a local optima. In addition, the rate of improvement decreases with the progression of the optimization. It is thus common that the best individual does not evolve over several generations when getting closer to the global solution. A value of $x=20$ generations was chosen.

@@

GAs are very computationally intensive because they require many evaluations of the objective function. These assessments are very long in this application, as they require calculating and assessing a forecast for every day of the calibration period, thus over several decades. In order to reduce the computation time, recalculating the performance score of an individual who has previously been evaluated and that has not changed was avoided. Thus, the score of each individual living in the selection was kept until it mutates.

@@

One of the goals being to make recommendations of parametrization in order to optimize the AM, a systematic procedure was adopted. The results are summarized hereafter \citep[see][for the details]{Horton2012a}. Concepts were taken from the factorial design approach \citep[see eg.][]{Costa2005a,Costa2007a,Mariano2010a}, which is sometimes used for comparative analysis of different parametrizations of GAs. It allows isolating the effect of a parameter under different combinations of the other options. A procedure by stages was adopted, analyzing in details and in a systematic way every variants of the implemented operators, in combination with multiple other options and parameters in order to take into account eventual co-dependencies. The goal here is not to focus on the performance score obtained through optimization, neither the values of the new optimized parameters \citep[covered in][]{Horton2016b}, but to explain how to use GAs to optimize AMs in an efficient way.

In order to evaluate a combination of operators/options, 10 optimizations per parametrization of GAs were processed. The performances were characterized by four indicators: (i) mean performance score: average of the final scores of the 10 optimizations, (ii) convergence: the number of optimizations that converged to a supposed global optimum, (iii) number of generations: characterization of the convergence speed, and (iv) number of evaluations of the objective function: characterization of the required calculation time.

@@

After a first overview of the results, GAs have quicky been proven successful at optimizing the AM. Indeed, the performance score of the sequential approach has been quickly exceeded without adding any new parameter to the method (Figure \ref{fig:gas_evolution_good}). Even GAs parametrization that will be further considered as inadequate (Figure \ref{fig:gas_evolution_bad}) did significantly better that the sequential approach.

@@

The results illustrate the effect of an operator when its contribution is isolated from the other operators. It means that we compare the effect of the studied operator when all the other operators are equivalent, while assessing multiple combinations. This contribution is then summarized relatively to the other options, as a percentage of gain/loss regarding the mean of all variants. For example, to evaluate the performance of the uniform crossover operator, its performance is compared to the average of all crossover operators while retaining the same population size, the same mutation operators, natural selection, and selection of couples.

@@

Every combination of 6 options for the couples selection (Table \ref{tab:assessed_couples_selection_operators}) and 21 for the chromosome crossover operators (Table \ref{tab:assessed_crossover_operators}) were evaluated, along with variants of the other operators. This results in 1,008 combinations, requiring 10,080 optimizations.

@@

The performance of the couples selection operator are relatively close (Figure \ref{fig:operator_selectcoupl_score}). Overall, the tournament selection with 3~candidates is slightly superior to others, along with the roulette wheel weighting. This last one is however a bit less effective in terms of convergence and number of evaluations (not shown). The couples selection operator has not a significant role in this application.

@@

Having identified the leading role of the mutation operator, the sensitivity analysis focused on it. Each of the 10 different implementations (see section \ref{sec:gas:mutation}) was tested, with different parameters for those who require some (Table \ref{tab:assessed_mutation_operators}), bringing the number of its variations up to 109. Some optimizations without any mutation were also performed as a reference. Along with variants of the other operators \citep[see][for the details]{Horton2012a}, it resulted in 660 combinations (so 6,600 optimizations).

@@

The mutation operators based on a variable normal or variable uniform laws work very poorly and are difficult to configure. Many operators present more or less with the same performance scores and require a variable amount of assessments. The convergence analysis \citep[see][]{Horton2012a} allows to highlight three best operators: non-uniform mutation, chromosome of adaptive search radius, and multi-scale mutation. Because of the importance of the mutation operator on the resulting performance, its role and parametrization were studied in more detail. Thus, different optimizations were performed using variants of these 3 operators (Table \ref{tab:assessed_mutation_operators_bis}).

@@

The first analysis was the optimization of the precipitation forecasting over a subcatchment (Binn-Simplon region) in the Swiss Alps (Figure \ref{fig:operator_mutation_score_atmlevel}). The optimizer could choose the 2 pressure levels of the atmospheric circulation analogy (method with a single level of analogy). The resulting performance score \citep[CRPSS = skill score of the CRPS, see][]{Brown1974} is obviously superior to the one from the sequential calibration (red line on Figure \ref{fig:operator_mutation_score_atmlevel}). For most options, it is also slightly better than the results from the optimization without selection of the pressure levels (blue line). A clear breakthrough of the performances was not expected, as the former selection of pressure levels results from intensive work \citep{Bontron2004}. This application however demonstrates that, when correctly parametrized, GAs can automatically and successfully choose the pressure levels. However, some less relevant optimizations do not match the previous results. Through different applications, the automatic selection of the pressure level was shown to significantly increase the difficulty for GAs to converge to a unique solution, ideally the global optimum. A difficulty is that the pressure levels are considered within the optimization without continuity between values, and thus the approaches relying on distance in the parameters space, such as the search radius, cannot fully exploit the properties that make them efficient. However, even though the results show a certain variability, most of them present very good performance scores, despite different parameters of the AM.

Then, the same optimization was performed, but for another region, sensitive to other meteorological influences (Figure \ref{fig:operator_mutation_score_rhoneamont}), in order to assess eventual dependencies of the operator with the predictand. Even though differences can be observed with Figure \ref{fig:operator_mutation_score_atmlevel}, it is globally the same options that perform better.

Next, a second level of analogy was added (Figure \ref{fig:operator_mutation_score_r2}) on moisture variables \citep[see][]{Bontron2004, Horton2016}. GAs had to optimize both levels of analogy simultaneously. Once again, despite the difficulty to optimize simultaneously both levels of analogy, the results were better than the sequential calibration (red line on Figure \ref{fig:operator_mutation_score_r2}). And finally, a preselection on air temperature was added instead of the fixed calendar window, as proposed by \cite{BenDaoud2015}. The results show generally higher scores (Figure \ref{fig:operator_mutation_score_r4}), demonstrating the success of the optimizer to take advantage of this new degree of freedom, and its capacity to handle optimization of 3 analogy levels jointly. Again, the most relevant options are generally the same.

After analysis of the most relevant mutation operators, the following advice can be raised (detailed parametrization are provided in section \ref{sec:recommendations}):

@@

	\item \textit{Non-uniform mutation} \citep{Michalewicz1996}: this operator is good in terms of convergence, mainly when the number of parameters to optimize is rather low. The number of required evaluations, however, can be quite substantial. The main disadvantage of the non-uniform mutation is the number of parameters it requires, that are difficult to estimate a priori. The $\omega$ coefficient does not influence performance. The role of $G_{m}$ is rather difficult to judge, but does not seem essential. The mutation rate was found to be important. The difficulty is that the optimal value may be case-related.

	\item \textit{Chromosome of adaptive search radius} (new): unlike the previous one, this new operator is very robust, as it requires no option and is auto-adapting. It is interesting to notice that the insertion of an extra chromosome representing the search radius gives better performance than other self-adaptive operators (such as, for example, the chromosome of adaptive mutation rate). If one had to choose a single option for the mutation operator, we would recommend this one, as it was proven effective and needs no parameter.

@@

In this application, the mutation operator has a leading effect and should be chosen with care. It may be wise to perform multiple optimizations and to consider these three operators in parallel in order to obtain results from options that are either sometimes more efficient or more robust. It is interesting to note that the three best techniques incorporate a notion of search distance. It is likely that this notion is the key to these algorithms, for this application, and allows them to initially explore the parameter domain, and then to converge. The search radius in fact directly represents the notion of transition between exploration and exploitation, in our opinion more than a possible evolution of mutation rates.

@@

The influence of the size of the IG (ratio of the total population) selected for mating was also assessed (Figure \ref{fig:option_popratio_score}). It does not appear that this parameter is critical to the quality of the optimizations, provided it is not too big. A value of 50~\% seems a wise choice.

@@

Optimizations by GAs on AMs of varying complexity were performed with a large number of combinations of operators in order to make recommendations for optimizing the AM. The conclusions are:

@@

The optimization does not systematically converge to the global optimum (but still often nearby), which is why it is recommended to do several optimizations in parallel in order to compare the results, analyze the convergence, and keep the best. It may be wise to consider the three mutation operators in parallel. It would then combine algorithms that are sometimes faster and other that are more robust. In order to be confident in the optimized AMs, we propose using a set of the following mutation operators:
